\mytitle{}

### 2018-11-14 paper/paper.tex

In service of this goal, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling for deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item A reactive simulation of web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention, improving upon past aggregate-based models (\cref{sec:acting-on-individual-sources}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

@@

This combination of algorithm and coding strategy is well-optimised.
If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
Action values for any state are then obtained by summation of the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}-1})$ floating point additions per decision.
Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
This means that we need only perform $n_{\mathit{tilings}} + 2$ additions and \num{2} multiplications per model update:
\begin{equation}
	\bm{w}_{t+1}[i][\operatorname{index}(A_t)] = \bm{w}_{t}[i][\operatorname{index}(A_t)] + \alpha \delta_t, \forall i \in s_{\mathit{list}}.
	\label{eqn:sg-sarsa-opt}
\end{equation}
If desired we may define a state-space with an arbitrary number of tiles per dimension (higher-resolution, lower generalisation), yet having constant-size state vectors and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
Beyond this, we need not store action values for tiles which have not yet been visited, conserving memory.

@@

We assume that the deployment environment is a moderately complex software-defined network, because the paradigm offers features which can directly benefit RL agents acting within.
The OpenFlow protocol allows a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allows relocation and easy installation of learners.

@@

At every time step, each agent installs an action via OpenFlow, instructing its host switch to drop each packet with probability $p$.
Specifically, the agent chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
In essence, this means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so exhibit a risk of introducing collateral damage into the network.
These choices primarily ensure that this rate reduction imposed on a source IP may never be performance.
It's worth noting that there are various ways that this could be done, and that the application of \emph{programmable data planes} to this end are suggested as future work.

@@

These are important issues, particularly when we consider real-world deployment.
Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most-used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network.

@@

\subsection{Service Guarantees and Host Density}
The default topology chooses $n=2$ hosts per egress point, which isn't a parameter setting that can be described as realistic for many web service deployments.
In practice, well-established network topologies such as the fat-tree topology fan out \emph{internally} (to provide efficient replication and scaling of resources) while having comparatively few egress points.
What this means is that datacentres and small-scale service deployments are more likely to have a host-to-learner ratio scaled heavily in favour of hosts, i.e. a large observed $n$.
We feel is most worth examining aspects of such a case, where the host/agent ratio is heavily skewed in favour of hosts and actions (as defined above) become less granular.

To test this, we configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single-server with a fan-out network.
The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
$U_s$ was fixed at $k \ell mn+2$ \si{\mega\bit\per\second}.
The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.

@@

From what we have seen, it is difficult (or impossible) for trace-based or numerical simulations to correctly capture certain dynamics without an extraordinary amount of care or consideration.
%As it turns out, 
Our goal is to briefly describe an environment which tests \emph{specific} behaviours to examine the \emph{specific} problems which have arisen during our testing of past approaches.
We contribute a network model built around live testing of TCP traffic in an SDN-enabled environment, adaptable to arbitrary topologies, with an explicit focus on preserving the real-time dynamics of TCP traffic.
In particular, we are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
Naturally, this model is not perfect or representative for all traffic, yet it captures some of the behaviour which we expect will plague most legitimate TCP flows.
If need be, we expect the frequency or distribution of requests could be conditioned to match observations of real-world access patterns.

@@

\subsection{Network Design}
We make use of a fully software-defined network, built using Openflow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
The controller is also responsible for computing how switches respond to ARP requests: the necessity of this arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.

\subsection{Traffic Model}
%?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
To model legitimate traffic, the server node runs an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
Each host's download rate is limited to match whatever maximum bandwidth was assigned to it.
This provides two modes of operation:
\begin{itemize}
	\item \emph{long flow} testing has each host constantly request a large binary file, such as a tarball of GCC8;
	\item \emph{request-chain} testing has each host request several random files known to exist within the website, followed by any dependent resources for each (stylesheets, images, etc.). On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime. This presents a more balanced distribution of flow duration and size.
\end{itemize}
We offer these to model two extremes: long-life host-to-server transfers (a characteristic we might e	xpect from e.g., streaming video), and common HTTP traffic.

%?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that simulation remains tractable.
When using TCP-like traffic, we must increase each attacker's upload range to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows---a \SI{30}{\percent} rate increase in expectation, uploading \SIrange{2.03}{2.18}{$\! \times U_s$}.
The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$

@@

%\fakepara{Timed Random Sequential actions}
Taking actions in this manner means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
This introduces numerous potential issues: inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated; at their worst, these present additional attack surfaces to an adversary.
To adapt to these problems, we make use of \emph{timed random sequential} updates.
Each agent begins with an empty work list.
For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
Uncompleted work is passed on to the next timestep, until it is emptied, at which point is is repopulated using the set of available measurements.
To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
Following \textcite{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 8 flows in this time.
We expect this should be tuned based on the frequency at which statistics arrive.
%?? Prune old? We can get through 8 per iteration, might use this as a basis for knowing when to say ``enough is enough''.

@@

%Explicitly mention \emph{why} we don't make use of neural networks, and instead favour more classical function approximation. Latency concerns, energy usage, graphics/TPU/whatever hardware availability or lack thereof across switches/routers/etc.
To make decisions cheaply and at low latency, we maintain the usage of semi-gradient Sarsa with tile coding, rather than making use of neural networks or more complicated function approximators.
Although this likely brings lower theoretical performance, there are more reasons than latency alone to favour classical methods; these include lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without the introduction of dedicated tensor processing hardware to networks.
Simpler decision boundaries reduce the risk of overfitting and of unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.

@@

This model is significantly different when considering how to tune the RL algorithm's parameters: each agent now requires the capability to plan ahead, since they may no longer choose $p$ directly from any state.
This entails choosing some discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
Recounting \cref{sec:reinforcement-learning}, the choice of $\gamma$ affects which fraction of the next current state-action pair's value is added to $R_t$ when updating the last state-action pair's value.
We found the setting  $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing with UDP traffic, by observing average performance over $\gamma$ increments of $0.1$.

@@

	\caption{
%		Features: how good are they really (part 2)? These are for TCP. These are less conclusive...
		Learned performance of Marl++ agents when benign traffic is TCP-like, using only a single feature as a basis for decisions (uncapped as in \cref{fig:udp-feature-plots}).
		Global state and Mean IAT still offer the greatest improvement above baseline, but it is observed that different optimal features are expressed between different traffic models.
		Packet-level statistics are considerably less effective for this class of traffic, with $\Delta$ Out Rate and Correspondence Ratio being proportionally better than they are alongside UDP.
		\label{fig:tcp-feature-plots}
	}
\end{figure}

@@

\fakepara{Mean packet size per-window}
Legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other classes of flooding attack.

@@

%?? NEED TO SAY SOMEWHERE: Baselines use TC for establishing interleaving/buffers filling etc., when MARL+ or SPF are in play then disabling TC (thus allowing illegal states, with \num{-1} reward marking them as such---``negative reinforcement'') may accelerate learning. We have a choice---but the results show that for TCP we WANT TC enabled... For UDP, we actually get a slight degradation in performance. This suggests a kind of ``banding'', perhaps? Important to mention that it easily learns how to maintain a logical bandwidth limit.
During training of the above feature models, we elected to remove the physical bandwidth limit to allow agents to visit illegal states and receive \emph{negative reinforcement} for doing so.
While this means that agents must learn how to keep traffic below a logical limit, they are helped by the fact that strong negative reward signals offer explicit information about which actions are poor choices: observe from \cref{eqn:reward,eqn:reward-but-better} that if a physical link has limit $U_s$ then we will never see a reward below 0.
It is important to note that neither setting affects the agents' ability to keep the total load below $U_s$.
%?? talk about cap difference: \cref{fig:tcp-cap-feature-plots}---we want cap enabled for TCP.
We found this led to a marginal improvement in efficacy when UDP was the primary source of benign traffic, but observed that this hindered learning when TCP was the primary protocol under protection (\cref{fig:tcp-cap-feature-plots}), likely due to the loss of key system dynamics.
It is anticipated that this difference might prove problematic for heterogeneous traffic mixtures.
One way of solving this may be to investigate the effect of replacing $U_s$ with e.g., $0.99 U_s$ in \cref{eqn:reward-but-better}, to introduce a small band of negative reinforcement without actually removing the load restriction in the environment.
It is crucial to make this region as small as possible, hence the somewhat arbitrary choice; we observe that agents are always able to learn how to keep traffic consistently below the logical bandwidth limit (as in the original work).
Our intent is to potentially accelerate learning while preserving the effects of system dynamics---we term these \emph{banded} rewards.

@@

Unless stated otherwise, these experiments measure average performance over 10 episodes, use all \num{14} features introduced in \cref{sec:rethinking-the-state-space} (tile-coded individually) alongside the global state, and make use of the physical bandwidth cap rather than pursuing negative reinforcement.
In all cases, agents learn control of the network from scratch.
Test system specifications match those given in \cref{sec:performance-in-an-emulated-environment}.

@@

	\caption{
		Online performance for UDP-like benign traffic.
		Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded) at small $n$, and remains marginally ahead of its predecessor by $n=16$, though both have undergone a significant drop in effectiveness.
		SPF, remarkably, displays performance on par with Marl++ for this more difficult topology in spite of underperforming when $n=2$.
		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
		The longer tails of outliers typically indicate the longer training time the new models require---we observe that SPF typically has considerably lower variance once it has converged on a stable policy.
		\label{fig:udp-box}
	}
\end{figure}

@@

	\caption{
		Online performance for TCP-like benign traffic.
		Marl and Marl++ begin with essentially identical efficacy, with Marl++ having a small advantage at larger $n$.
		SPF is slightly worse than Marl and Marl++ at low $n$, but achieves considerably better median performance and lower variance than the other models for higher $n$.
		\label{fig:tcp-box}
	}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{../plots/tcp-16-single}

	\caption{
		Online performance for $n=16$ hosts per egress point when benign traffic is TCP-like.
		At this level of host density, SPF reaches its peak performance sooner and is considerably more consistent throughout.
		Marl++ takes longer to reach its most effective policy, we achieve visibly better peak performance.
		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
		\label{fig:tcp-16}
	}
\end{figure}

@@

\subsection{Banded rewards and negative reinforcement}
In general, we find that explicitly training a policy to keep the system in a valid state via negative reinforcement (i.e., an \emph{uncapped} network) is most effective for UDP-like traffic with Marl++.
For SPF, however, negative reinforcement is typically a worse choice.
Where TCP-like traffic is concerned in both models, we see that this strategy almost always leads to worse policies.

The usefulness of banded rewards with regard to learning an effective policy appears to be unpredictable.
For many of the configurations detailed in \cref{tab:av-vals}, for $n>2$ the average reward with banding is likely to lie between that of the capped and uncapped system models, but oscillates heavily.
In a handful of UDP instances, banding can lead to \emph{better} average performance.
More study is likely required into this; we choose to replace $U_s$ with $0.99U_s$ to minimise the impact on legitimate states, but this choice remains somewhat arbitrary.

>>>>

In service of this goal, this paper contributes:
\begin{itemize}
%	\item A DDoS mitigation system based on direct-control reinforcement learning, designed for deployment in real-world software-defined networks (\cref{sec:environment-and-rl-algorithm,sec:acting-on-individual-sources}).
	\item Important weaknesses and flaws in the past design and evaluation of similar techniques \cite{DBLP:journals/eaai/MalialisK15}. We offer a deconstruction and empirical study on their formulation, flaws, and risk factors concerning different traffic distributions---in addition, we demonstrate modifications enabling deployment in real-world software-defined networks (\cref{sec:performance-in-an-emulated-environment}).
	\item A reactive simulation of web-server traffic, designed to test system characteristics that packet trace playback fails to capture (\cref{sec:a-new-normal}).
	\item An in-depth investigation into suitable features for an automatic DDoS mitigation system, with qualitative and quantitative justification (\cref{sec:rethinking-the-state-space}).
	\item Two source-level granularity approaches to RL-driven DDoS prevention, improving upon past aggregate-based models (\cref{sec:acting-on-individual-sources}), alongside an empirical evaluation (\cref{sec:the-results-of-doing-so}).
\end{itemize}

@@

This combination of algorithm and coding strategy is well-optimised.
If $\operatorname{A}(s) = \mathcal{A}$ in every state (and actions are themselves discrete), these properties allow a particularly efficient (vectorised) implementation of the policy and update rules by storing a vector of action values for each state encountered.
Action values for any state are then obtained by summation of the weight vectors from all activated tiles---taking $|\mathcal{A}|(n_{\mathit{tilings}}-1)$ floating point additions per decision.
Observing that $\nabla{\acval{s}{a}{\wvec{}}} = \operatorname{\mathbf{x}}(s, a)$, further optimisations arise by considering that a tile-coded feature vector is a binary vector of constant Hamming weight (and so is amenable to representation as an array of indices, $s_{\mathit{list}}$).
This means that we need only perform $n_{\mathit{tilings}} + 2$ additions and \num{2} multiplications per model update:
\begin{equation}
	\bm{w}_{t+1}[i][\operatorname{index}(A_t)] = \bm{w}_{t}[i][\operatorname{index}(A_t)] + \alpha \delta_t, \forall i \in s_{\mathit{list}}.
	\label{eqn:sg-sarsa-opt}
\end{equation}
If desired we may define a state-space with an arbitrary number of tiles per dimension (higher-resolution, lower generalisation), yet having constant-size state vectors and constant action computation cost ($\mathcal{O}(n_{\mathit{tilings}})$).
Beyond this, we need not store action values for tiles which have not yet been visited, conserving memory.

@@

We assume that the deployment environment is a moderately complex software-defined network, because the paradigm offers features which can directly benefit RL agents acting within.
The OpenFlow protocol allows a controller (or other authorised hosts) to install complex actions, forwarding rules and logic into a switch at runtime.
Furthermore, networks of this kind more naturally enable the future use of \emph{network function virtualisation}, a technology which could allow relocation and easy installation of learners.

@@

At every timestep, each agent installs an action via OpenFlow, instructing its host switch to drop each packet with probability $p$.
Specifically, the agent chooses $p \in \left\{ 0.0, 0.1, \ldots, 0.9 \right\}$, giving a discrete, static action set which cannot completely filter traffic.
These choices primarily ensure that the rate reduction imposed on a source IP may never be permanent or irreversible.
In essence, this means that each agent is in control of pushback \cite{DBLP:journals/ccr/MahajanBFIPS02a}, and so carries a risk of introducing collateral damage into the network.
It's worth noting that there are various ways that this could be implemented, and that the application of \emph{programmable data planes} to this end are suggested as future work.

@@

These are important issues, particularly when we consider real-world deployment.
Heuristic estimates of traffic legitimacy come with computational cost and couple the reward function to the accuracy of the estimator, hosts often show diversity in their own traffic patterns (perhaps being multi-modal), and it is known that TCP is the most used transport protocol for Internet traffic \cite{DBLP:conf/saint/ZhangDJC09}.
Different network environments will also impose different levels of host density, where popular web servers may have orders of magnitude more clients than egress points from their network.

@@

\subsection{Service Guarantees and Host Density}
The default topology chooses $n=2$ hosts per egress point, which isn't a parameter setting that can be described as realistic for many web service deployments.
In practice, well-established network topologies such as the fat-tree topology fan out \emph{internally} (to provide efficient replication and scaling of resources) while having comparatively few egress points.
What this means is that datacentres and small-scale service deployments are more likely to have a host-to-learner ratio scaled heavily in favour of hosts, i.e. a large observed $n$.
We feel is most worth examining aspects of such a case, where actions (as defined above) become less granular.

To test this, we configured the network topology using $k=2$ teams, $\ell=3$ intermediate nodes per team, $m=2$ agents per intermediate node, and $n \in \{2, 4, 8, 16\}$ hosts per learner.
This is a slight simplification of the \textquote{online} experiment as presented by \textcite{DBLP:journals/eaai/MalialisK15}, choosing fewer teams but remaining as a single server with a fan-out network.
The algorithm parameters were set at $\gamma=0$ (leading to opportunistic behaviour), $\alpha=0.05$, having linearly annealed $\epsilon=0.2 \rightarrow 0$ by $t=3000$.
Benign and malicious hosts uploaded between \SIrange{0}{1}{\mega\bit\per\second} and \SIrange{2.5}{6}{\mega\bit\per\second} respectively, and hosts were redrawn at each episode's start with $\operatorname{P}(\mathit{malicious})=0.4$.
$U_s$ was fixed at $k \ell mn+2$ \si{\mega\bit\per\second}.
The performance of each choice of $n$ was averaged over \num{10} episodes of length \num{10000} timesteps (setting each agent's $\wvec{}=\bm{0}$ between episodes).
Host allocations were generated pseudorandomly to ensure fairness between choices of $n$.
These parameter choices match those of the original study to enable direct comparison, and are (to the best of our knowledge) arbitrary, but we justify our range of $n$ as capturing increasing scales of host activity.

@@

From what we have seen, it is difficult (or impossible) for trace-based or numerical simulations to correctly capture certain dynamics without an extraordinary amount of care or consideration.
%As it turns out, 
Our goal is to briefly describe an environment which tests \emph{specific} behaviours to examine the \emph{specific} problems which have arisen during our testing of past approaches.
We contribute a network model built around live testing of TCP traffic in an SDN-enabled environment, which is adaptable to arbitrary topologies, with an explicit focus on preserving the real-time dynamics of TCP traffic.
In particular, we are interested in capturing interactive, correlated back-and-forth exchanges associated with live HTTP traffic; mainly because of the particular interactions between the application-level dynamics, congestion awareness at the transport level and the nature of control signal used.
Naturally, this model is not perfect or representative for all traffic, yet it captures some of the behaviour which we expect will plague most legitimate TCP flows.
If need be, we expect the frequency or distribution of requests could be conditioned to match observations of real-world access patterns.

@@

\subsection{Network Design}
We make use of a fully software-defined network, built using OpenFlow-aware switches in mininet alongside a controller application based on \emph{Ryu} \cite{ryu}.
All internal routers are primed with knowledge of the shortest path to each internal host, while new inbound flows register the ``way back'' for each hop used, to ensure consistent traffic conditions for each flow.
If such information is lost, perhaps expiring due to inactivity, it suffices to forward an outbound packet on a random (outbound) port, as we assume that any external IP is reachable through any of the test network's egress ports (i.e., that it is not connected to any stub autonomous systems).
The controller is also responsible for computing how switches respond to ARP requests: the necessity of this arises due to the reliance upon Linux's networking stack for live applications, and wouldn't need to be considered for purely trace-based evaluation.
We make further use of the topology presented earlier (\cref{sec:topology}), noting that our architecture allows us to trivially extend and modify this if required.

\subsection{Traffic Model}
%?? Legitimate traffic: TCP traffic (HTTP clients downloading web pages, dependent resources and files) with a mixture of lifetimes for each request.
To model legitimate traffic, the server node runs an nginx v1.10.3 HTTP daemon, serving statically generated web pages alongside various large files and binaries.
Benign hosts run a simple libcurl-based application written in Rust, repeatedly requesting resources from the server.
Each host's download rate is limited to match whatever maximum bandwidth was assigned to it.
This provides two modes of operation:
\begin{itemize}
	\item \emph{long flow} testing has each host constantly request a large binary file, such as a tarball of GCC8;
	\item \emph{request-chain} testing has each host request several random files known to exist within the website, followed by any dependent resources for each (stylesheets, images, etc.). On completion, a host changes its IP to generate separate statistics per-flow, while minimising downtime. This presents a more balanced distribution of flow duration and size.
\end{itemize}
We offer these to model two extremes: long-life host-to-server transfers (a characteristic we might expect from e.g., streaming video), and common HTTP traffic.

%?? Malicious traffic: UDP flood traffic (hping3, MTU-size packets, ). Why not min-size packets? Because the traffic generator gets in a horrible rut if I do so...
Malicious traffic is generated by use of the \emph{hping3} program, generating UDP-flood traffic targeting random ports.
Each malicious packet is MTU-sized, loosely mapping to the large frame sizes expected in an amplification attack and to ensure that emulation remains tractable.
When using TCP-like traffic, we must increase each attacker's upload range to \SIrange{4}{7}{\mega\bit\per\second} to meaningfully impact benign flows---a \SI{30}{\percent} rate increase in expectation, uploading \SIrange{2.03}{2.18}{$\! \times U_s$}.
The expected fraction of $U_s$ consumed by each host is \SI{21.5}{\percent} for $n=2$, and \SI{2.84}{\percent} for $n=16$.

@@

%\fakepara{Timed Random Sequential actions}
Taking actions in this manner means that any agents are assigned a larger, and potentially unbounded, set of tasks to perform every time they receive load and flow statistics from the network and their parent switch.
This introduces numerous potential issues: inability to respond to unexpected changes in flow state, delayed service of new flows, and risks that flow states become outdated; at their worst, these present additional attack surfaces to an adversary.
To adapt to these problems, we make use of \emph{timed random sequential} updates.
Each agent begins with an empty work list.
For the set of flows active in any timestep, we shuffle the list and perform as many action calculations and updates as possible, within a set time limit.
Uncompleted work is passed on to the next timestep, until the list is emptied, at which point is is repopulated using the set of available measurements.
To ensure that flow control actions are made with recent information, we combine state vectors for unvisited flows in the current work set, and replace the stored vector for all others.
State vector combination is done by summing deltas and packet counts, updating means via weighted sums, and replacing all other fields with the newest measurements.
Following \citeauthor{DBLP:conf/sigcomm/ChenL0L18}'s observations concerning short flows \cite{DBLP:conf/sigcomm/ChenL0L18}, we maintain a deadline of \SI{1}{\milli\second}---in tests, an agent is typically able to process around 8 flows in this time.
We expect this should be tuned based on the frequency at which statistics arrive.
%?? Prune old? We can get through 8 per iteration, might use this as a basis for knowing when to say ``enough is enough''.

@@

%Explicitly mention \emph{why} we don't make use of neural networks, and instead favour more classical function approximation. Latency concerns, energy usage, graphics/TPU/whatever hardware availability or lack thereof across switches/routers/etc.
To make decisions cheaply and at low latency, we maintain the usage of semi-gradient Sarsa with tile coding, rather than making use of neural networks or more complicated function approximators.
Although this likely brings lower theoretical performance, there are more reasons than latency alone to favour classical methods; these include lower energy usage, reduced model complexity (and training time), the availability of necessary hardware, and simpler decision boundaries.
This aligns with our goal of quick online learning, and potentially faster adaptation to aggregate changes in traffic without necessarily introducing dedicated tensor processing hardware to networks.
Simpler decision boundaries reduce the risk of overfitting and of unexpected behaviour, and we anticipate that the simplicity of policy computation in the chosen manner will also greatly aid interpretability of anomalous action choices.

@@

This model is significantly different when considering how to tune the RL algorithm's parameters: each agent now requires the capability to plan ahead, since they may no longer choose $p$ directly from any state.
This entails choosing some discount factor $\gamma \ne 0$, allowing the value of future states to influence state-action value updates.
Recounting \cref{sec:reinforcement-learning}, the choice of $\gamma$ affects which fraction of the current state-action pair's value is added to $R_t$ when updating the value of the last state-action pair.
We found the setting $\gamma = 0.8$ to be the most effective choice for this hyperparameter during exploratory testing with UDP traffic, by observing average performance over $\gamma$ increments of $0.1$.

@@

	\caption{
%		Features: how good are they really (part 2)? These are for TCP. These are less conclusive...
		Learned performance of Marl++ agents when benign traffic is TCP-like, using only a single feature as a basis for decisions (uncapped as in \cref{fig:udp-feature-plots}).
		Global state and Mean IAT still offer the greatest improvement above baseline, but it is observed that different optimal features are expressed between different traffic models.
		Packet-level statistics are considerably less effective for this class of traffic, with $\Delta$ Out Rate and Correspondence Ratio being proportionally better than they are for UDP.
		\label{fig:tcp-feature-plots}
	}
\end{figure}

@@

\fakepara{Mean packet size per window}
Legitimate flows, both TCP- and UDP-based, often transmit packets with a distribution of sizes.
Attack traffic is not likely to be so diverse: we might expect solely max-size packets in the case of amplification attacks, or minimum-size packets in other classes of flooding attack.

@@

%?? NEED TO SAY SOMEWHERE: Baselines use TC for establishing interleaving/buffers filling etc., when MARL+ or SPF are in play then disabling TC (thus allowing illegal states, with \num{-1} reward marking them as such---``negative reinforcement'') may accelerate learning. We have a choice---but the results show that for TCP we WANT TC enabled... For UDP, we actually get a slight degradation in performance. This suggests a kind of ``banding'', perhaps? Important to mention that it easily learns how to maintain a logical bandwidth limit.
During training of the above feature models, we elected to remove the physical bandwidth limit to allow agents to visit illegal states and receive \emph{negative reinforcement} for doing so.
While this means that agents must learn how to keep traffic below a logical limit, they are helped by the fact that strong negative reward signals offer explicit information about which actions are poor choices: observe from \cref{eqn:reward,eqn:reward-but-better} that if a physical link has limit $U_s$ then we will never see a reward below 0.
It is important to note that neither setting affects the agents' ability to keep the total load below $U_s$.
%?? talk about cap difference: \cref{fig:tcp-cap-feature-plots}---we want cap enabled for TCP.
We found this led to a marginal improvement in efficacy when UDP was the primary source of benign traffic, but observed that this hindered learning when TCP was the primary protocol under protection (\cref{fig:tcp-cap-feature-plots}), likely due to the loss of key system dynamics.
It is anticipated that this difference might prove problematic for heterogeneous traffic mixtures.
One way of solving this may be to investigate the effect of replacing $U_s$ with e.g., $0.99 U_s$ in \cref{eqn:reward-but-better}, to introduce a small band of negative reinforcement without actually removing the load restriction in the environment.
It is crucial to make this region as small as possible, hence the somewhat arbitrary choice of factor; we observe that agents are always able to learn how to keep traffic consistently below the logical bandwidth limit (as in the original work).
Our intent is to potentially accelerate learning while preserving the effects of system dynamics---we term these \emph{banded} rewards.

@@

Unless stated otherwise, these experiments measure average performance over 10 episodes, use all \num{14} features introduced in \cref{sec:rethinking-the-state-space} (tile-coded individually) alongside the global state, and make use of the physical bandwidth cap rather than pursuing negative reinforcement.
In all cases, agents learn control of the network from scratch ($\wvec{}=\bm{0}$).
Test system specifications match those given in \cref{sec:performance-in-an-emulated-environment}.

@@

	\caption{
		Online performance for UDP-like benign traffic.
		Marl++ offers a marked improvement (a peak $\sim$\SI{30}{\percent} more benign traffic arrives unimpeded) at small $n$, and remains marginally ahead of its predecessor by $n=16$, though both have undergone a significant drop in effectiveness.
		SPF, remarkably, slightly outperforms Marl++ (with lower variance) for this more difficult topology despite being worse for smaller $n$.
		Both of the new models take longer to train, but achieve better peak and average performance than Marl.
		The longer tails of outliers typically indicate the longer training time the new models require---we observe that SPF typically has considerably lower variance once it has converged on a stable policy.
		\label{fig:udp-box}
	}
\end{figure}

@@

	\caption{
		Online performance for TCP-like benign traffic.
		Marl and Marl++ begin with essentially identical efficacy, with Marl++ having a small advantage at larger $n$.
		SPF is slightly worse than Marl and Marl++ at low $n$, but achieves considerably better median performance and lower variance than the other models for large $n$.
		\label{fig:tcp-box}
	}
\end{figure}
\begin{figure}
	\centering
	\includegraphics[width=0.9\linewidth]{../plots/tcp-16-single}

	\caption{
		Online performance of standard and single-agent models with $n=16$ hosts per egress point and TCP-like benign traffic.
		At this level of host density, SPF reaches its peak performance sooner and is considerably more consistent throughout the episode.
		Marl++ takes longer to reach its most effective policy, and achieves visibly better peak performance than the other standard models.
		With a single agent, Marl++ shows worse performance, while SPF improves significantly and continues to learn well past annealing $\epsilon \rightarrow 0$.
		\label{fig:tcp-16}
	}
\end{figure}

@@

\subsection{Banded rewards and negative reinforcement}
In general, we find that explicitly training a policy to keep the system in a valid state via negative reinforcement (i.e., an \emph{uncapped} network) is most effective for UDP-like traffic with Marl++.
For SPF, however, negative reinforcement is typically a worse choice.
Where TCP-like traffic is concerned in both models, we see that this strategy almost always leads to worse policies and higher collateral damage.

The usefulness of banded rewards with regard to learning an effective policy appears to be unpredictable.
For many of the configurations detailed in \cref{tab:av-vals}, for $n>2$ the average reward with banding is likely to lie between that of the capped and uncapped system models, but oscillates heavily.
In a handful of UDP instances, banding can lead to \emph{better} average performance.
More study is likely required into this; we replace $U_s$ with $0.99U_s$ to minimise the impact on legitimate states, but this choice remains somewhat arbitrary.
